---
title: "Mid-semester check-in: Part I"
format: gfm
---

1. Calculate the average and standard deviation of flight delays for each airline, *counting only those flights that were delayed longer than an hour*. 

```{r message=FALSE}
library(nycflights13)
library(tidyverse)

flights_b <- group_by(flights, carrier) |>
  filter(!is.na(arr_delay),
         arr_delay > 60)

summary_flights <- summarize(flights_b,
                             mean_arr_delay = mean(arr_delay),
                             sd_arr_delay = sd(arr_delay)
                             )

slice_head(ungroup(summary_flights), n=10)
```

2. Identify the 5 flights (i.e., unique combination of `carrier` and `flight`) that are most frequently delayed more than a half hour. 

```{r message=FALSE}

flights_c <- group_by(flights, carrier, flight) |>
  filter(!is.na(arr_delay),
         arr_delay > 30)

summary_flights_c <- summarize(flights_c, n = n()) |>
  arrange(desc(n))

slice_head(ungroup(summary_flights_c), n=5)
```

3. Make a plot of the distribution of flight times for each airline. The plot should show summary statistics in some way (e.g. median, upper and lower quartiles; you can choose which summary statistics are best) as well as individual data points. Order the x axis in order of median flight time, from highest on the left to lowest on the right.

```{r message=FALSE}

flights_d <- group_by(flights, carrier) |>
  filter(!is.na(air_time))

ggplot(flights_d, 
       aes(x = fct_reorder(.f = carrier, .x = air_time, .fun = median, .desc = TRUE), 
         y = air_time)) +
  geom_boxplot(outlier.alpha = 0.1, outlier.size=0.1) +
  geom_jitter(data=slice_sample(flights_d, prop=0.3),
              color="blue", size=0.1, alpha=0.1)
```

